# 1 · Getting financial data {.hero #s01-get-data}

::: {.section-badge}
Module 1 · Import (R4DS ch 7)
:::

Loading market data

::: {.lede}
Two paths: pull from Yahoo Finance live, or read the CSVs shipped with this course.
:::

## The classic way: `tidyquant::tq_get()`

Live, from Yahoo Finance:

```r
library(tidyquant)

aapl <- tq_get("AAPL",
               from = "2015-01-01",
               to   = "2024-12-31",
               get  = "stock.prices")
head(aapl)
```

Fine for exploratory use, but renders need the network. For a reproducible course, we cache to disk once and read from CSV.

## The reproducible way: committed CSVs

```{r}
#| message: false
library(readr)
library(dplyr)

aapl <- read_csv("data/AAPL.csv", show_col_types = FALSE)
head(aapl, 4)
```

Every subsequent module reads from `data/` — no network calls at render time.

## The five tickers you'll see

- **AAPL** — Apple, large-cap equity
- **MSFT** — Microsoft, another large-cap equity (for correlation stories)
- **SPY** — S&P 500 ETF, the market proxy for CAPM
- **TLT** — 20+ Year Treasury Bond ETF, bond proxy
- **GLD** — SPDR Gold Shares, commodity proxy

## Fama-French factors, too

```{r}
#| message: false
ff <- read_csv("data/ff_factors.csv", show_col_types = FALSE)
head(ff, 3)
```

`Mkt-RF`, `SMB`, `HML`, `RF` in decimals (not percent). Used in module 9 for CAPM.
