# 2 · Visualize with ggplot2 {.hero #s02-first-plot}

::: {.section-badge}
Module 2 · Visualize (R4DS ch 2)
:::

Your first plot

::: {.lede}
`ggplot()` builds plots as layers: data + aesthetics, then geoms, then labels.
:::

## Just draw the line

```{r}
#| message: false
#| fig-align: center
#| fig-width: 8
#| fig-height: 4
library(readr); library(dplyr); library(ggplot2)

aapl <- read_csv("data/AAPL.csv", show_col_types = FALSE)

ggplot(aapl, aes(date, adjusted)) +
  geom_line(colour = "#325D88") +
  labs(title = "AAPL adjusted close", x = NULL, y = "USD") +
  theme_minimal()
```

That's the entire grammar: **data** (`aapl`), **aesthetics** (`aes(date, adjusted)`), **geom** (`geom_line`), **labs**, **theme**.

## Three tickers on the same axes

```{r}
#| message: false
#| fig-align: center
#| fig-width: 8
#| fig-height: 4
library(purrr)

tickers <- c("AAPL", "SPY", "GLD")
prices <- map_dfr(tickers,
                  ~ read_csv(paste0("data/", .x, ".csv"), show_col_types = FALSE) |>
                     mutate(ticker = .x))

ggplot(prices, aes(date, adjusted, colour = ticker)) +
  geom_line() +
  scale_colour_manual(values = c(AAPL = "#325D88", SPY = "#B94A48", GLD = "#557A3E")) +
  labs(title = "Ten years of daily close", x = NULL, y = "Adjusted close (USD)", colour = "Ticker") +
  theme_minimal()
```

## Now the deconstructed animation

The slide below plays out the same idea layer-by-layer:

{{< include sections/demo-ggplot.qmd >}}
