# 7 · Write your own functions {.hero #s07-functions}

::: {.section-badge}
Module 7 · Program (R4DS ch 25)
:::

Wrap the analysis

::: {.lede}
When you find yourself pasting the same expression three times, it's a function. Finance is full of repeated formulas — Sharpe, drawdown, annualisation. Learn to write them yourself before reaching for `PerformanceAnalytics`.
:::

## Anatomy

```{r}
sharpe_ratio <- function(returns, rf = 0, freq = 252) {
  excess <- returns - rf / freq
  mean(excess, na.rm = TRUE) / sd(excess, na.rm = TRUE) * sqrt(freq)
}

# Try it on AAPL daily log returns
library(readr); library(dplyr)
aapl_ret <- read_csv("data/AAPL.csv", show_col_types = FALSE) |>
  mutate(r = log(adjusted / lag(adjusted))) |>
  pull(r)

sharpe_ratio(aapl_ret) |> round(3)
```

## Maximum drawdown

```{r}
max_drawdown <- function(returns) {
  cum   <- cumprod(1 + returns[!is.na(returns)])
  peaks <- cummax(cum)
  min((cum - peaks) / peaks)
}

max_drawdown(aapl_ret) |> round(3)
```

Interpretation: the largest peak-to-trough loss you'd have experienced holding AAPL over the full sample.

## Compare to a library implementation

```{r}
#| warning: false
#| message: false
library(PerformanceAnalytics)
# PerformanceAnalytics expects an xts; convert quickly
aapl_xts <- read_csv("data/AAPL.csv", show_col_types = FALSE) |>
  mutate(r = log(adjusted / lag(adjusted))) |>
  filter(!is.na(r)) |>
  { \(d) xts::xts(d$r, order.by = d$date) }()

SharpeRatio.annualized(aapl_xts, Rf = 0) |> round(3)
maxDrawdown(aapl_xts) |> round(3)
```

The numbers should match your homemade functions to within rounding.
