# Vectors & Lists {.hero #s07-vectors-lists}

::: {.section-badge}
Module 7
:::

Vectors & Lists

::: {.lede}
Atomic vectors hold one type. Lists hold anything. That's the whole distinction.
:::

## Building atomic vectors {data-anim="code-type"}

```{r}
x <- c(1, 2, 3, 4, 5)
x
typeof(x)
length(x)
```

`c()` is the workhorse constructor: it **c**ombines values into a vector.

## What `c()` actually does {data-anim="vector-fill"}

::: {.r-stack data-anim="vector-fill" data-vector="[1,2,3,4,5]"}
:::

::: {.muted style="text-align:center; margin-top:1em;"}
Each cell holds one value; the order is preserved; the type is shared.
:::

## Coercion rules {data-anim="code-type"}

```{r}
# Numbers + logical → numeric (TRUE=1, FALSE=0)
c(1, 2, TRUE)

# Anything + character → all character
c(1, "two", TRUE)

# The hierarchy: logical → integer → double → character
```

When `c()` sees mixed types, it promotes everything to the most-permissive type.

## Lists hold anything {data-anim="code-type"}

```{r}
fit_summary <- list(
  model    = "lm",
  coef     = c(intercept = 0.4, slope = 1.2),
  fitted   = c(0.4, 1.6, 2.8, 4.0),
  formula  = "y ~ x"
)
fit_summary$coef
```

Lists are how R returns complex objects: `lm()`, `summary()`, almost every modelling function returns a list.

---

::: {.section-badge}
Recap
:::

## Module 7 · Recap {data-anim="stagger-up"}

- `c()` builds atomic vectors; one type per vector
- Coercion follows logical → integer → double → character
- Lists are the catch-all heterogeneous container; `$name` accesses elements
