# Matrices & Arrays {.hero #s08-matrices-arrays}

::: {.section-badge}
Module 8
:::

Matrices & Arrays

::: {.lede}
A matrix is a 2-D atomic vector with a `dim()` attribute. Arrays add more dimensions.
:::

## Constructing a matrix {data-anim="code-type"}

```{r}
m <- matrix(1:12, nrow = 3, ncol = 4, byrow = TRUE)
m
dim(m)
```

`matrix()` reshapes a vector. `byrow = TRUE` fills row-by-row;
`byrow = FALSE` (the default) fills column-by-column.

## Visualizing the fill order {data-anim="matrix-grid"}

::: {.r-stack data-anim="matrix-grid" data-matrix="[[1,2,3,4],[5,6,7,8],[9,10,11,12]]"}
:::

::: {.muted style="text-align:center; margin-top:1em;"}
Watch the cells fill row-by-row, just as `byrow = TRUE` instructed.
:::

## Naming dimensions {data-anim="code-type"}

```{r}
rownames(m) <- c("alpha", "beta", "gamma")
colnames(m) <- c("Q1", "Q2", "Q3", "Q4")
m

# Then you can index by name:
m["alpha", "Q3"]
```

For higher dimensions, use `array(values, dim = c(...))`. You'll meet 3-D
arrays mainly in image processing and certain Bayesian outputs.

---

::: {.section-badge}
Recap
:::

## Module 8 · Recap {data-anim="stagger-up"}

- `matrix(vec, nrow, ncol)` — 2-D, single type, ordered
- `byrow` controls fill order; `dim()` reports shape
- `rownames`/`colnames` enable name-based indexing
