# 9 · Linear regression: CAPM {.hero #s09-capm}

::: {.section-badge}
Module 9 · Model (R4DS light)
:::

Alpha and beta

::: {.lede}
The Capital Asset Pricing Model in one line: `lm(r_asset - rf ~ r_mkt - rf)`. Slope is beta, intercept is alpha, R² is how much of the asset's variance the market explains.
:::

## Load the pieces

```{r}
#| message: false
library(readr); library(dplyr); library(broom); library(lubridate)

# AAPL excess returns
aapl <- read_csv("data/AAPL.csv", show_col_types = FALSE) |>
  mutate(r_aapl = log(adjusted / lag(adjusted))) |>
  select(date, r_aapl)

ff <- read_csv("data/ff_factors.csv", show_col_types = FALSE) |>
  select(date, mkt_rf, rf)

capm_data <- inner_join(aapl, ff, by = "date") |>
  mutate(r_aapl_excess = r_aapl - rf) |>
  filter(!is.na(r_aapl_excess))

head(capm_data, 3)
```

## Fit the regression

```{r}
fit <- lm(r_aapl_excess ~ mkt_rf, data = capm_data)
summary(fit)
```

## Read the coefficients

```{r}
tidy(fit)
glance(fit) |> select(r.squared, adj.r.squared, sigma, statistic, p.value)
```

- **Intercept** (`alpha`) — the "excess" return AAPL earns per day beyond what CAPM predicts. Statistically indistinguishable from zero here.
- **Slope** (`beta`) — how much AAPL moves for each 1% move in the excess market return. β ≈ 1.2 means AAPL is somewhat more volatile than the market.
- **R²** — fraction of AAPL's daily variance explained by the market. ~0.4 for a large-cap tech name is typical.

## Extend to a factor model (three factors)

```{r}
ff_full <- read_csv("data/ff_factors.csv", show_col_types = FALSE)
capm3 <- inner_join(aapl, ff_full, by = "date") |>
  mutate(r_aapl_excess = r_aapl - rf) |>
  filter(!is.na(r_aapl_excess))

lm(r_aapl_excess ~ mkt_rf + smb + hml, data = capm3) |> tidy()
```

Same one-line change adds `SMB` (size) and `HML` (value) — the Fama-French 3-factor model.
