# Data Frames & Tibbles {.hero #s09-dataframes-tibbles}

::: {.section-badge}
Module 9
:::

Data Frames & Tibbles

::: {.lede}
Each column is a variable, each row is an observation. The everyday R table.
:::

## Meet the penguins {data-anim="code-type"}

```{r}
library(palmerpenguins)
library(tibble)

penguins
```

The `penguins` dataset is a **tibble**: 344 rows × 8 columns of measurements
on three species of penguin from the Palmer Archipelago, Antarctica.

## `data.frame` vs `tibble` {data-anim="stagger-up"}

::: {.columns}

::: {.column width="50%"}
**`data.frame`** (base R, since 1990s)

- Prints **all** rows by default
- Coerces strings to factors (historically)
- `df[, "col"]` may return a vector or a frame
:::

::: {.column width="50%"}
**`tibble`** (tidyverse, modern)

- Prints **10 rows** + column types
- Strings stay strings, no factor surprise
- `tb[, "col"]` always returns a tibble
:::

:::

```{r}
class(penguins)
```

A tibble *is* a data frame (it inherits from `data.frame`), with friendlier
defaults — anything that takes a `data.frame` takes a tibble.

## Column access {data-anim="code-type"}

```{r}
# Single column as a vector
head(penguins$bill_length_mm, 5)

# Subset of columns as a tibble
penguins[, c("species", "bill_length_mm")]
```

The `$` operator is the most common way to grab a single column.

## Quick summary {data-anim="code-type"}

```{r}
dim(penguins)
names(penguins)
summary(penguins$body_mass_g)
```

`summary()` adapts to the column type: numeric columns get five-number
summaries; factors get counts.

---

::: {.section-badge}
Recap
:::

## Module 9 · Recap {data-anim="stagger-up"}

- Each column = one variable; each row = one observation
- Tibbles print friendlier and behave more predictably than base `data.frame`
- `$col`, `summary()`, and `head()` are your three first calls on any new frame
