# 10 · Communicate with Quarto {.hero #s10-communicate}

::: {.section-badge}
Module 10 · Communicate (R4DS ch 29)
:::

## The minimum

Create `report.qmd`:

````r
---
title: "AAPL: a mini portfolio note"
author: "Gabriel Cabrera"
format: html
execute:
  echo: true
  warning: false
---

## Setup

```{r}
library(readr); library(dplyr); library(ggplot2)
aapl <- read_csv("data/AAPL.csv", show_col_types = FALSE)
```

## Price history

```{r}
ggplot(aapl, aes(date, adjusted)) + geom_line() + theme_minimal()
```

## Summary

- Mean daily log return: `{r} round(mean(log(aapl$adjusted / lag(aapl$adjusted)), na.rm = TRUE), 5)`
````

Then in the terminal:

```
quarto render report.qmd
```

## Publication-grade tables with `gt`

```{r}
#| message: false
library(gt)

capm_results <- tibble(
  ticker = c("AAPL", "MSFT", "SPY", "GLD"),
  alpha  = c(0.0002, 0.0001, 0.0000, 0.0001),
  beta   = c(1.21, 1.09, 1.00, 0.06),
  r2     = c(0.42, 0.51, 1.00, 0.02)
)

capm_results |>
  gt() |>
  tab_header(title = "CAPM regression summary") |>
  fmt_number(columns = c(alpha), decimals = 4) |>
  fmt_number(columns = c(beta, r2), decimals = 2) |>
  cols_label(ticker = "Ticker", alpha = "α", beta = "β", r2 = "R²")
```

## Ship the whole analysis

Pick your output format in the YAML:

```yaml
format: revealjs      # for a slide deck like this one
format: html          # for a notebook-style report
format: pdf           # for a printable PDF
format: docx          # for a Word document
```

Same `.qmd`, four different outputs. Add `execute: freeze: auto` to cache chunk results so successive renders only re-run edited chunks.
